Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion

### 2023-09-14 Manuscript.Rmd
editing for PAIN

---
title: "Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion"
author: Alexandra G. Mitchell^1^,
        Jesper Fischer Ehmsen^1^,
        Daniel Elmstrøm Christensen^1^,
        Anna Villaume Stuckert^2^,
        Patrick Haggard^3^,
        Francesca Fardo^1,4^
output:
  word_document:
      reference_docx: knitting_files/test_markdown.docx
  html_document:
    css: knitting_files/style.css
  pdf_document:
    latex_engine: xelatex
indent: true
bibliography: knitting_files/tgi-spine2.bib
csl: knitting_files/nature.csl
link-citations: yes
linkcolor: blue
serif: Times New Roman
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: sentence
---

@@

# Results and discussion

@@

While most previous studies have focused on the TGI as a single experience, explicitly understood as an experience of or like pain, our approach differed.
We instructed participants to report the sensation from just one of the thermal components, corresponding to the cold thermode location (Exp. 1) or the warm thermode location (Exp. 2).
Importantly, the aim of this method was not to isolate any possible independent effect of warm and cold stimulation during TGI, but rather to ensure consistency in participants' ratings across trials and different stimulus conditions.

To investigate both the thermosensory and burning components in each experiment, we asked participants to provide a subjective evaluation of multiple sensory qualities - cold, warmth and burning (pain).
Stimuli consisted of either cold-warm pairs (TGI stimuli), which potentially evoked an illusion of heat and pain, or control stimuli that involved pairing a cold (Exp. 1) or a warm (Exp. 2) stimulus of identical temperature as in the TGI condition with a baseline temperature of 30°C (non-TGI stimuli).
All stimulation pairs were presented at a fixed distance on the skin, either within the same dermatome or across dermatomes that mapped onto non-adjacent spinal segments (Figure 1A and B).
After ten seconds of stimulation, participants reported their ratings using three sequential VAS scales ranging from 0, indicating the lack of the corresponding sensory quality, to 100, indicating an extreme sensation (Figure 1C), presented in a random order.
For each instance of stimulation, participants were guided to focus their reporting on the sensations originating either thermode 'A' or 'B' which corresponded to either the cold (Exp. 1) or warm (Exp. 2) thermode.
This labelling was used to ensure participants were not immediately aware of the temperature associated with each probe.
VAS ratings for each sensation were analysed using mixed-effects zero-inflated beta regressions, where VAS ratings that equaled zero for non-TGI and TGI stimulation were modelled separately to VAS ratings above zero.

@@

## Thermosensory and burning components of TGI perception are spinally mediated

@@

## Spinal organisation mediates the TGI

>>>>

---
title: "Disentangling the spinal mechanisms of illusory heat and burning sensations in the Thermal Grill Illusion"
author: Alexandra G. Mitchell^1^,
        Jesper Fischer Ehmsen^1^,
        Daniel Elmstrøm Christensen^1^,
        Anna Villaume Stuckert^2^,
        Patrick Haggard^3^,
        Francesca Fardo^1,4^
output:
  word_document:
      reference_docx: knitting_files/test_markdown.docx
  html_document:
    css: knitting_files/style.css
  pdf_document:
    latex_engine: xelatex
indent: true
bibliography: knitting_files/tgi-spine2.bib
csl: knitting_files/ieee.csl
link-citations: yes
linkcolor: blue
serif: Times New Roman
always_allow_html: yes
editor_options: 
  markdown: 
    wrap: sentence
---

@@

@@

While most previous studies have focused on the TGI as a single experience, explicitly understood as an experience of or like pain, our approach differed.
We instructed participants to report the sensation from just one of the thermal components, corresponding to the cold thermode location (Exp. 1) or the warm thermode location (Exp. 2).
Importantly, the aim of this method was not to isolate any possible independent effect of warm and cold stimulation during TGI, but rather to ensure consistency in participants' ratings across trials and different stimulus conditions.To investigate both the thermosensory and burning components in each experiment, we asked participants to provide a subjective evaluation of multiple sensory qualities - cold, warmth and burning (pain).
Stimuli consisted of either cold-warm pairs (TGI stimuli), which potentially evoked an illusion of heat and pain, or control stimuli that involved pairing a cold (Exp. 1) or a warm (Exp. 2) stimulus of identical temperature as in the TGI condition with a baseline temperature of 30°C (non-TGI stimuli).
All stimulation pairs were presented at a fixed distance on the skin, either within the same dermatome or across dermatomes that mapped onto non-adjacent spinal segments (Figure 1A and B).

@@

## Procedure

# Results
## Thermosensory and burning components of TGI perception are spinally mediated

@@

# Discussion
